---
title: "Lab_6"
author: "Adam Hayes, Jasmine Sanchez, Erin Omyer, Richard Park"
date: "2/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
COflights <- read_csv("https://files.osf.io/v1/resources/3z5as4pej8/providers/osfstorage/5a8ca28f57103100104584db")


```

###The shorter the flight distance the increase in arrival delay
```{r}

ggplot(data = COflights) + geom_point(mapping = aes(x = DISTANCE, y = ARR_DELAY), color = 'blue', position = 'dodge', na.rm = TRUE) + coord_flip()

```
#When the distance of a flight decreases, the chances of an arrival delay are higher. The longer the flight distance, the less chance of an arrival delay. This means that a short flight distance increases the arrival delay.

###Erin's Findings

#Probability of delayed flights to Denver with air time greater than 4 hours
```{r}
prob_delayDEN <- COflights %>%
  filter(ARR_DELAY > 0 & AIR_TIME > 240, DEST == 'DEN') %>%
  count()
total_delayDEN <- COflights %>%
  filter(ARR_DELAY > 0, DEST == 'DEN') %>%
  count()

prob_delayDEN/total_delayDEN
```
#Probability of delayed flights to Denver with air time less than 2 hours
```{r}
prob_delayDEN2 <- COflights %>%
  filter(ARR_DELAY > 5 & AIR_TIME < 120, DEST == 'DEN') %>%
  count()
total_delayDEN <- COflights %>%
  filter(ARR_DELAY > 0, DEST == 'DEN') %>%
  count()
prob_delayDEN2/total_delayDEN
```
#From the probabilities displayed above, as the air time increases, the probability of an arrival delay decreases. As the air time decreases, then the probability of an arrival delay increases. It is roughly around 4 percent probabilty of a delayed arrival for a flight longer than 4 hours. But, for flights less than 2 hour, there is a 40 percent probability of a delayed arrival. 
